---
title: "Class10"
author: "Gen Dantay BIMM 143"
format: pdf
---

# Cooperative analysis of ADK

ADK (Adenalte Kinase) is an important drug target and we would love to know how it works - i.e. molecular mechanism 

There has been lots of work done on this protein due to it's importance including lots of crystal structures. 

We will begin with getting an example ADK sequence from the database. We will then use this to find all ADK structures in the PDB

```{r}
library(bio3d)
aa <- get.seq("1ake_A")
aa
```

We can now run BLAST with this sequence 

```{r}
#b <- blast.pdb(aa)
```

```{r}
#hits <- plot(b)
```

Let's see what is in our `hits` object 

```{r}
#hits$pdb.id

hits<- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```

Now we can download all these PDB structure files:
```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```
```{r}
files
```


Now I want to align and superimpose these structures which are all over the place. 

```{r}
# align related PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

Let's have a look at our `pdbs` object

```{r}
pdbs
```

Now we have our aligned and superposed structures, we can perform all sorts of analysis on them. Let's do PCA...

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```

Results of PCA on Adenylate Kinase X-ray structures. Each dot represents the PDB structure.

We can cluster the structures by RMSD (or any other method)

```{r}
# Calculate RMSD
rd<- rmsd(pdbs)
```
```{r}
# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

We can make a movie - also called a trajectory of the major differences (i.e. structural displacements) of ADK 

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

We can also plot our main PCA results with ggplot 

```{r}
#Plotting results with ggplot2
library(ggplot2)
library(ggrepel)
ids <- basename.pdb(pdbs$id)

df <- data.frame(PC1=pc.xray$z[,1], 
                 PC2=pc.xray$z[,2], 
                 col=as.factor(grps.rd),
                 ids=ids)

p <- ggplot(df) + 
  aes(PC1, PC2, col=col, label=ids) +
  geom_point(size=2) +
  geom_text_repel(max.overlaps = 20) +
  theme(legend.position = "none")
p

```


